```{r}
pacman::p_load(
  tidyverse,
  skimr, janitor,# exploracion de los datos
  lme4, ordinal, gamlss,  # modelado
  ggResidpanel,  # diagnostico 
  DHARMa, 
  performance,   # evaluar performance de los modelos
  emmeans,       # medias estimadas por el modelo  
  multcomp       # comparar las medias entre si - tukey
  )
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
```

```{r}
# https://www2.gov.bc.ca/gov/content/industry/agriculture-seafood/animals-and-crops/plant-health/insects-and-plant-diseases/tree-fruits/fire-blight

vinc_raw <- read_csv("https://github.com/juanchiem/agro_data/raw/master/SummerInoculationFinal.csv") %>% 
  janitor::clean_names() %>% 
  select(year, apogee, proca=apogee_bino, pruned, plot, score_factor) 

# vinc_raw %>% rio::export(here::here("glm2/vinc.csv"))
vinc_raw <- rio::import(here::here("glm2/vinc.csv"))
vinc_raw
vinc <- vinc_raw %>% 
  mutate_at(vars(score_factor, year, plot, apogee, proca, pruned), as.factor) %>% 
  mutate(score_factor=ordered(score_factor, 
                              levels = c("0","1","2","3","4")), 
         pruned = fct_rev(pruned))
vinc$score_factor
```

```{r}
str(vinc)
glimpse(vinc)
vinc %>% tabyl(year, apogee, pruned)
```

```{r}
mod0 <- clmm(score_factor ~ 1 + (1|plot), data=vinc)

mod1 <- update(mod0, ~ . + apogee * year * pruned)
anova(mod0, mod1)
print(summary(mod1),digits=1)

mod2 <- update(mod1, ~ year*proca + year*pruned   +(1|plot))
anova(mod1, mod2)
compare_performance(mod0,mod1, mod2)

print(summary(mod2),digits=1)
# comparar con Table 4
sjPlot::tab_model(mod2)
```

The overall disease score was higher in 2019 than 2018 (LR=191 (1 df); P<0.001), presumably in part because disease had more time to progress in shoots between inoculation and observation (7 days in 2018 and 9 days in 2019). 

```{r}
# year2019    2.6        0.2      11   <2e-16 ***

mod_year <- update(mod0, ~ . + year)
anova(mod0, mod_year)
```

The disease score of shoots from plots sprayed with ProCa was lower (CLMM: LR=19.5 (1 df); P<0.001) than for the unsprayed,

```{r}
# procaYes              -1.5        0.3      -4    2e-05 ***
mod_proca <- update(mod0, ~ . + proca)
anova(mod0, mod_proca)
```

and the ProCa efect was greater in 2019 (interaction) (CLMM: LR=6.3 (1 df); P=0.011). 

```{r}
mod_proca_adit <- update(mod0, ~ . + proca+year)
mod_proca_int <- update(mod0, ~ . + proca*year)
anova(mod_proca_adit, mod_proca_int)
# year2019:procaYes     -0.6        0.3      -2     0.02 *  
```

The pruning efect was diferent between years (interaction) (CLMM: LR=21.9 (2 df); P<0.001).  Summer pruning reduced the disease score in 2018 but had no efect in 2019. 

```{r}
mod_pruned_adit <- update(mod0, ~ . + pruned+year)
mod_pruned_int <- update(mod0, ~ . + pruned*year)
anova(mod_pruned_adit, mod_pruned_int)

# year2019:prunedNo  -1.1        0.2      -5    4e-06 ***
# > emm_mod1. = emmeans(mod2, ~   pruned | year)
# > cld(emm_mod1.)
# year = 2018:
#  pruned emmean    SE  df asymp.LCL asymp.UCL .group
#  Yes      1.06 0.274 Inf     0.524      1.60  1    
#  No       1.85 0.276 Inf     1.312      2.39   2   
# 
# year = 2019:
#  pruned emmean    SE  df asymp.LCL asymp.UCL .group
#  No       3.07 0.281 Inf     2.518      3.62  1    
#  Yes      3.35 0.282 Inf     2.798      3.90  1    

```

There was no interaction between ProCa sprays and pruning (CLMM: LR=7.6 (1 df); P=0.27) or higher order interaction. 

```{r}
mod_procapruned_adit <- update(mod0, ~ . + pruned + proca)
mod_procapruned_int <- update(mod0, ~ . + pruned*proca)
anova(mod_procapruned_adit, mod_procapruned_int)

# mod1
# apogeeNone:prunedNo                 0.07       0.79     0.1     0.93    
# apogeeSymptom:prunedNo              0.58       0.80     0.7     0.47 
```
The efect of ProCa (odds ratio (OR) approx. 4× in 2018 and 8× in 2019) was the same for both spray initiation timings (CLMM: LR=1.8 (2 df); P=0.41).
```{r}
# anova(mod1, mod2)
# mod1     16 2537.1 -1252.6  5.4586  4     0.2434
```


```{r}
emm_mod1 = emmeans(mod2, ~ proca * pruned | year)
cld(emm_mod1)

emm_mod1. = emmeans(mod2, ~   pruned | year)
cld(emm_mod1.)
# pred <- ggpredict(mod.py.yab, c("pruned", "apogee_bino"))
```

```{r}
vinc %>% 
  mutate(apogee=fct_relevel(apogee, "None")) %>% 
  ggplot()+
  aes(x = pruned, fill = score_factor)+
  geom_bar(position = position_fill(reverse = TRUE))+
  facet_grid(year ~ apogee )+
  labs(x="", y="Proportion of shoots")+
  scale_fill_grey(name = 'Score', start =.89, end = 0 )+
  scale_x_discrete(labels = c("Pruned", "Control"))
```
